A Low Complexity Adaptive Beamformer\\for Active Sonar Imaging

### 2012-06-18 HRUIonGPU/jrnl_article.tex

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

@@

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

@@

@@

@@

@@

>>>>

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

@@

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
%\let\MYorigsubfloat\subfloat
%\renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

@@

(Go in detail about the implementation using sliding DFT).

@@

\subsection{Toeplitz-based solver} 
(cite new article). Discuss it.

@@

\begin{figure*}[!t]
\centerline{\subfloat[]{\includegraphics[width=3.5in]{gfx/benchmark_us_32.png} \label{fig:benchUS32}}
\hfil
\subfloat[]{\includegraphics[width=3.5in]{gfx/benchmark_us_64.png}%
\label{fig:benchUS64}}}
\caption{Benchmark of GPU Capon beamforming for a cardiac image covering a 70 degrees sector. Execution time is plotted as a function of subarray length $L$. The plot shows execution times for the three steps listed in Section \ref{sec:meth}. Calculating covariance matrices in red, solver in green, and the final beamformer sum in blue. For each step, there is one line for three different choices of temporal averaging, $N_{avg} \in [0, 1, 2]$. (a) A 32 elements array. (b) A 64 elements array.}
\label{fig:bench}
\end{figure*}

@@

How has the data been acquired?

### 2012-06-18 UAM2011/Article/article2.tex

\documentclass[
   UAM                                          % Document type (non-standard)
 , 12pt                                         % Text size
%  , draft
%  , final                                        % Quality
 , xelatex                                      % Use the XeLaTeX compiler
%  , biblatex                                     % Use 'biblatex' for references (best by far)
 , bibtex                                       % Use 'bibtex' for references (oldschool :)
%  , movie
%  , notodos                                      % Disable todos
 , layout
%  , defaultformat                                % Attempt to use default formatting options
%  , glossary                                     % Use a glossary
]{common/mytemplate}

@@

% \setlength\floatsep{\baselineskip}
\setcounter{topnumber}{1}
% \setlength\baselineskip{12pt}
% \selectfont
\begin{document}

\pagestyle{plain}
% \layout
% \printinunitsof{cm}
% \prntlen{\baselineskip}
% \the\textwidth
% \the\linewidth
% \prntlen{\textwidth}\\
% \prntlen\linewidth\\
% \prntlen\oddsidemargin\\
% \prntlen\oddsidemargin
% \newpage

\title{A Low Complexity Adaptive Beamformer\\for Active Sonar Imaging}%
%
\author{J.I.Buskenes\firstAddress, A.Austeng\firstAddress, C.-I.C.Nilsen\firstAddress}%
%
\begin{contact}
  \firstAddress Dept. of Informatics, P.Box 1080 Blindern, N-0316 OSLO
\end{contact}%
%
\begin{contact}
Contact author: Jo Inge Buskenes\\
Dept. of Informatics, Univ. of Oslo, P.Box 1080 Blindern, N-0316 OSLO\\
\href{mailto:joibu@ifi.uio.no}{joibu@ifi.uio.no}
\end{contact}%
%
\begin{abstract}
The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels.

While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output.

@@

An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA method is of O($MW$), with $M$ being the number of channels and $W$ the number of windows. We made the LCA perform like the MV method using a well designed set of 30 windows. Hence, unless the array is very small, the proposed method will perform like the MV beamformer or better, and at a fraction of the computational cost.
\end{abstract}%
%
\keywords{Beamforming, Capon, Minimum variance, Low Complexity, Sonar}

@@

In active sonar array imaging, an acoustic wave is transmitted and the received echoes are usually recorded using an array of sensors. Each of these may be independently delayed and weighted such that signals emanating from directions of interest are added constructively, while noise and interference from other directions add destructively. Additionally, a window may be applied to the sensor channels to further adjust the arrays spatial response. This process is known as beamforming.% The process of combining the signals from each sensor is known as beamforming.

Traditional beamformers such as Delay-and-Sum (DAS) weight the sensors using a predefined window. However, since the optimal spatial response is likely to change with time, it is often preferred to use beamformers that continously adapt the weights to the dynamic data. This usually involves solving some optimization problem.

The Minimum Variance (MV) beamformer selects the window that minimizes the beamformer's output power~\cite{cap69}. It has recently been investigated for active sonar imaging in e.g.~\cite{saf09}, and in ultrasound imaging in e.g.~\cite{syn07}. We suggest using a Low Complexity Adaptive (LCA) beamformer, which for each pixel applies a set of predefined windows and selects one of them based on the MV criterion. Using the LCA beamformer results in increased robustness and improved edge definition compared to the MV and DAS beamformers.

@@

The MV beamformer computes the window coefficients by estimating and inverting a spatial covariance matrix. For improved estimation accuracy the data is averaged in space or time, or both. In addition, a certain amount of diagonal loading is often added to the covariance matrix before inversion. These actions, while leading to statistical robustness, tend to deteriorate the beamformer's performance. Also, the inversion of the covariance matrix has a computational complexity of O($M^3$), where $M$ is the number of elements. The computational burden makes the MV beamformer for large arrays less attractive, and sometimes even infeasible.

Based on the proposed method by Synnevåg~\cite{syn11}, we have implemented the LCA beamformer that keeps the minimum variance optimisation criterion but reduces the solution space to a discrete set of predefined windows. This reduces the computational complexity to O($MW$), where $W$ is the number of windows in the set. We use the Kaiser window function because it allows us to design a wide range of windows with different mainlobe widths and sidelobe suppression by adjusting the tradeoff parameter $\beta$~\cite{kai66}. In addition we apply steering to each of the windows and constrain the window design to ensure unit gain in the look direction.

@@

\begin{figure}[tp]\centering%
\subfigure{%
\graphicsAI[width=\linewidth]{gfx/img_das_lca_mv.pdf}}
\subfigure{%
\graphicsAI[width=\linewidth]{gfx/img_das_lca_mv_mean100.pdf}}%
\caption{A simulated circular object buried in speckle is imaged from above.\newline\parbox[t]{1.50cm}{Top:\newline Bottom:}\parbox[t]{6.8cm}{Image of a single speckle realisation.\newline Mean image of 100 speckle realisations.}}\label{speckle}%%
\end{figure}%
\begin{figure}[t]\centering%
\graphicsAI[width=\linewidth]{gfx/img_mean_cut.pdf}%
\caption{Lateral cut in the mean image at range 41.1\,m (highlight) and 44.5\,m (shadow).}\label{cutmean}%
\end{figure}
% \begin{figure}[t]\centering
% % \graphicsAI[drawing,width=\linewidth]{gfx/the_box.svg}\\
% \includegraphics[width=\linewidth]{gfx/selected_windows.pdf}
% \caption{Selected windows}\label{selected_windows}
% \end{figure}
% \begin{figure}[t]\centering
% \includegraphics[width=\linewidth]{gfx/selected_windows_angle.pdf}
% \caption{Selected windows - Colored by angle}\label{selected_windows_angle}
% \includegraphics[width=\linewidth]{gfx/selected_windows_sym_angle.pdf}
% \caption{Selected windows - Colored symmetrically by angle}\label{selected_windows_sym}
% \end{figure}
\begin{figure}[t]\centering
\graphicsAI[width=\linewidth]{gfx/img_holmengraa.pdf}%
\caption{HISAS sidescan sonar (SSS) image of the shipwreck Holmengraa.}\label{holmengraa}%
\end{figure}

% Adaptive beamformers usually compute the window coefficients by estimating and inverting a spatial covariance matrix. There are two inherent problems to this procedure. First, for improved estimation accuracy the data is either averaged in space or time, and the significance of the covariance matrix can be reduced using a regularisation parameter~\cite{Carl Inge}. These are all attempts to constrain the beamformer to ensure the window coefficience are not over-adapted to the data. Second, inverting the covariance matrix has a computational complexity of O($M^3$)~\cite{Carl Inge}. For larger arrays the computational burden becomes significant, and the question arises whether this processing power can be better utilised. 
% 
% Based on the proposed method by Synnevåg~\cite{syn11}, we have implemented a Low Complexity Adaptive (LCA) beamformer that keeps the minimum variance optimisation criteria but reduces the solution space to a discrete set of predefined windows. We use the Kaiser window function because it allows us to design a wide range of windows with different mainlobe widths and sidelobe suppression by adjusting the tradeoff parameter $\beta$. In addition we apply steering $\varphi$ to each of the windows and and constrain the window design to ensure unit gain in the look direction.

@@

For the LCA to perform well it is important to let it have a diverse selection of windows to choose from. This selection is based on observations of the solutions found by the MV beamformer. We obtained good results in our experiments by letting the LCA beamformer select among 5 Kaiser windows with uniformly distributed $\beta$'s in the range $[0.05, 0.5]$ and a rectangular window. Each window was steered in 5 different directions, uniformly distributed within 80\% of the mainlobe width of the rectangular window, which is the most narrow. This adds to a total of 30 windows.

To make the choice of window less susceptible to pixel value uncertainties, the LCA beamformer was set to apply the most frequently selected window in an 11 pixel range region to the center pixel in that region. In physical terms, this is a region of approximately 20\,cm. The MV beamformer estimated the covariance matrix as described in~\cite{syn07} by averaging over 16 subarrays and 11 range pixels, and applying 3\% diagonal loading.

@@

We have tested the beamformers using data aquired by the Kongsberg Maritime HISAS1030 sonar~\cite{han09}, and on simulations of the same sonar with 100 realisations of speckle.

The simulations contained a few highlighted cylindric regions with 2\,m diameter and a constant intensity 15.4\,dB over the average speckle level. We focused on an object centered at 41\,m range and -3 degrees azimuth angle. The images produced by the LCA, MV and unweighted DAS for a single speckle realisation and a mean image of all 100 realisations are shown in \Fig{speckle}. Each image was normalised by their respective speckle level, and the dynamic range was clamped at \{-30,~15\}\,dB. The physical extent of the of the simulated object was superimposed for reference. %\ref{saf09}

In \Fig{speckle} we observe that both the adaptive beamformers produce images with a more clearly defined shadow. The edges are sharper and the shadow more distinct. This is because the adaptive beamformers have better sidelobe suppression, and thus allow less signal energy to leak from the speckle region into the shadow region. The same effects can also be observed laterally around the highlight region of the mean image; DAS is more prone to smear energy into the speckle region than the adaptive beamformers are.

\Fig{cutmean} displays two lateral cuts in the mean image, one through the highlight at 41.1\,m range and one through the shadow at 44.5\,m range. Each cut was computed as the mean of a 40\,cm range band centered at the respective ranges. The transition region between highlight and shadow is shortest for the LCA, which is advantageous since it results in a more accurate representation of the object size. The slightly inferior performance of the MV beamformer is due to the subarray averaging, which makes the effective array smaller~\cite{syn07}.

By studying the LCA performance on the 100 speckle realisations, we found the choice of window to be highly dependent on the phenomena being imaged. In speckle regions LCA favored narrow spatial responses with minor steering, while wide and fully steered responses were often selected in shadow regions. There was also a clear distinction between the responses selected in sidelobe regions and in speckle regions.

@@

A sidescan image of the 1500 dwt oil tanker wreck Holmengraa is shown in \Fig{holmengraa}. It is about 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth~\cite{holmengraa}. The sidescan image was created using data from the HISAS 1030 sonar, which is rather unsuited for this purpose because of its large opening angle. This, and the fact that the wreck was imaged at a range of about 105\,m makes the image quality poor, but sufficient to compare our beamformers.

In the Holmengraa image we note again that the LCA produces a cleaner shadow and better edge definition. The MV method performed almost identically to LCA in this case, and was omitted.

@@

The LCA method performs similar to the MV method, but at a fraction of the computational cost. The key to achieve this success lies in the design of the window set. Windows that yield vagely steered narrow responses are preferred in speckle regions, while wide and steered responses are typically preferred in highlight and shadow regions. A window set of 6 different responses each steered in 5 different directions proved sufficient in our experiment to match the performance of the MV beamformer.

The tendency of the LCA to perform similarly to the MV beamformer seems to indicate that a full solution space is rarely needed in real world scenarios.

% Nature, while being stochastic, acts within predictable bounds\todo{hehe, jada, jeg skal fjerne..} that can be well adapted to by using beamformers with fewer degrees of freedom than what the MV method provides.

% While the MV beamformer is expected to take advantage of its full solution space to design a spatial response with zero sensitivity at the interferers angles, while LCA will be limted to the spatial responses that we selected. In short, whenever a scenario comes along that favors a spatial response we did not think of, the MV will triumph. On the other hand, increasing the LCA solution space will at some point leave it with nearly the same freedom as the MV, without suffering from statistical instability.
% \todo[inline]{mer innhold..}

% While e found that the tendency of the LCA to perform similarly to the MV beamformer is consistent in most scenarios, except when it is of essence that strong interferers close to object being imaged are present\todo{referanse?}. In such cases the MV beamformer is expected to take advantage of its full solution space to design a spatial response with zero sensitivity at the interferers angles, while LCA will be limted to the spatial responses that we selected. In short, whenever a scenario comes along that favors a spatial response we did not think of, the MV will triumph. On the other hand, increasing the LCA solution space will at some point leave it with nearly the same freedom as the MV, without suffering from statistical instability.
% \todo[inline]{mer innhold..}

\newpage
\bibliography{references}

>>>>

\documentclass[
   UAM                                          % Document type (non-standard)
%  , draft
%  , final                                        % Quality
%  , xelatex                                      % Use the XeLaTeX compiler
%  , biblatex                                     % Use 'biblatex' for references (best by far)
 , bibtex                                       % Use 'bibtex' for references (oldschool :)
%  , movie
%  , notodos                                      % Disable todos
 , layout
%  , defaultformat                                % Attempt to use default formatting options
%  , glossary                                     % Use a glossary
]{common/mytemplate}

@@

% \setlength\floatsep{\baselineskip}
\setcounter{topnumber}{1}
% \setlength\baselineskip{12pt}
% \selectfont
\begin{document}
% \fontsize{12pt}{12pt}\selectfont

\layout
\pagestyle{plain}
% \layout
% \printinunitsof{cm}
% \prntlen{\baselineskip}
% \the\textwidth
% \the\linewidth
% \prntlen{\textwidth}\\
% \prntlen\linewidth\\
% \prntlen\oddsidemargin\\
% \prntlen\oddsidemargin
% \newpage

\title{A Low Complexity Adaptive Beamformer\\for Active Sonar Imaging}%
%
\author{J.I.Buskenes\firstAddress, A.Austeng\firstAddress, C.-I.C.Nilsen\firstAddress}%
%
\begin{contact}
  \firstAddress Dept. of Informatics, P.Box 1080 Blindern, N-0316 OSLO
\end{contact}%
%
\begin{contact}
Jo Inge Buskenes\\
Dept. of Informatics, Univ. of Oslo, P.Box 1080 Blindern, N-0316 OSLO\\
\href{mailto:joibu@ifi.uio.no}{joibu@ifi.uio.no}
\end{contact}%
%
\begin{abstract}
The angular resolution and contrast in active sonar images depend on the beamformer's ability to receive signals from directions of interest, while suppressing noise and interference emanating from other directions. For sonar arrays, this is achieved by applying weights to the array channels. While classical beamformers use predefined windows, adaptive beamformers estimate the optimal window by analytical evaluation of the data. The minimum variance (MV) beamformer, for instance, calculates the set of weights that minimises the variance of the beamformer's output.

@@

An attactive trait of the LCA is its low computational complexity; while the MV beamformer is of O($M^3$), the LCA is of O($MW$). Here $M$ is the number of channels, $W$ the number of windows. Hence the proposed method performs like a MV beamformer or better, but at a fraction of the computational cost.
\end{abstract}%
%
\keywords{Beamforming, Capon, Minimum variance, Low Complexity, Sonar}

@@

In active sonar array imaging, an acoustic wave is transmitted and the received echoes are recorded using an array of sensors. Each of these may be independently delayed and weighted such that signals emanating from directions of interest are added constructively, while noise and interference from other directions add destructively. Additionally, a window may be applied to the sensor channels to further adjust the arrays spatial response. This process is known as beamforming.% The process of combining the signals from each sensor is known as beamforming.

Traditional beamformers such as Delay-and-Sum (DAS) weight the sensors using a predefined window. Adaptive beamformers find an optimal window for each pixel in the resulting image. These windows are usually found by solving an optimization problem based on the data.

The Minimum Variance (MV) beamformer selects the window that minimizes the beamformer's output power~\cite{cap69}. It has previously been investigated for active sonar imaging in e.g. \cite{saf09}. We suggest a Low Complexity Adaptive (LCA) beamformer, which for each pixel applies a set of predefined windows and selects one of them based on the MV criterion. Using the LCA beamformer results in increased robustness and improved edge definition compared to the MV and DAS beamformers.

% 959 27 818

@@

The MV beamformer computes the window coefficients by estimating and inverting a spatial covariance matrix. For improved estimation accuracy the data is averaged in space or time, or both. In addition, a certain amount of diagonal loading is often added to the covariance matrix before inversion. These actions, while leading to statistical robustness, often also deteriorate the beamformers performance. Also, the inversion of the covariance matrix has a computational complexity of O($M^3$), where $M$ is the number of elements. The computational burden makes the MV beamformer for large arrays less attractive, and sometimes even infeasible.

Based on the proposed method by Synnevaag~\cite{syn08}, we have implemented the LCA beamformer that keeps the minimum variance
optimisation criterion but reduces the solution space to a discrete set of predefined windows. This reduces the computational complexity to O($MW$), where $W$ is the number of windows in the set. We use the Kaiser window function because it allows us to design a wide range of windows with different mainlobe widths and sidelobe suppression by adjusting the tradeoff parameter $\beta$. In addition we apply steering to each of the windows and constrain the window design to ensure unit gain in the look direction.

@@

\begin{figure}[tp]\centering%
\subfigure{%
\graphicsAI[width=\linewidth]{gfx/img_das_lca_mv.pdf}%
}
\subfigure{
\graphicsAI[width=\linewidth]{gfx/img_das_lca_mv_mean100.pdf}%
}\label{speckle}%
% \caption{\parbox[t]{1.50cm}{Top:\protect\newline Bottom:}\parbox[t]{6.8cm}{Image of a single speckle realisation.\protect\newline Mean image of 100 speckle realisations.}}%
\end{figure}%
\begin{figure}[t]\centering%
\graphicsAI[width=\linewidth]{gfx/img_mean_cut.pdf}%
\caption{Cut through shadow and highlight region of the mean image.}\label{cutmean}%
\end{figure}
% \begin{figure}[t]\centering
% % \graphicsAI[drawing,width=\linewidth]{gfx/the_box.svg}\\
% \includegraphics[width=\linewidth]{gfx/selected_windows.pdf}
% \caption{Selected windows}\label{selected_windows}
% \end{figure}
% \begin{figure}[t]\centering
% \includegraphics[width=\linewidth]{gfx/selected_windows_angle.pdf}
% \caption{Selected windows - Colored by angle}\label{selected_windows_angle}
% \includegraphics[width=\linewidth]{gfx/selected_windows_sym_angle.pdf}
% \caption{Selected windows - Colored symmetrically by angle}\label{selected_windows_sym}
% \end{figure}
\begin{figure}[t]\centering
\graphicsAI[width=\linewidth]{gfx/img_holmengraa.pdf}%
\caption{HISAS Synthetic Aperture Sonar (SAS) image of the shipwreck Holmengraa}\label{holmengraa}%
\end{figure}

% Adaptive beamformers usually compute the window coefficients by estimating and inverting a spatial covariance matrix. There are two inherent problems to this procedure. First, for improved estimation accuracy the data is either averaged in space or time, and the significance of the covariance matrix can be reduced using a regularisation parameter~\cite{Carl Inge}. These are all attempts to constrain the beamformer to ensure the window coefficience are not over-adapted to the data. Second, inverting the covariance matrix has a computational complexity of O($M^3$)~\cite{Carl Inge}. For larger arrays the computational burden becomes significant, and the question arises whether this processing power can be better utilised. 
% 
% Based on the proposed method by Synnevåg~\cite{syn08}, we have implemented a Low Complexity Adaptive (LCA) beamformer that keeps the minimum variance optimisation criteria but reduces the solution space to a discrete set of predefined windows. We use the Kaiser window function because it allows us to design a wide range of windows with different mainlobe widths and sidelobe suppression by adjusting the tradeoff parameter $\beta$. In addition we apply steering $\varphi$ to each of the windows and and constrain the window design to ensure unit gain in the look direction.

@@

For the LCA to perform well it is important to let it have a diverse selection of windows to choose from. This selection is based on observations of the solutions found by the MV beamformer. We obtained good results in these experiments by letting the LCA beamformer select among 10 Kaiser windows with uniformly distributed parameters ($\beta$) in the range $[0.05, 0.5]$ and a rectangular window. Each window is steered in 10 different directions, uniformly distributed within 80\% of the mainlobe width caused by the rectangular window, which is the most narrow. This adds to a total of 110 windows.

For statistical robustness the LCA beamformer was set to apply the most frequently selected window in an 11 pixel range region to the center pixel in that region. The MV beamformer estimated the covariance matrix as described in~\cite{syn08} by averaging over 16 subarrays and 11 range pixels, and applying 3\% diagonal loading.

@@

% We have tested the beamformers using data aquired by the Kongsberg Maritime HISAS1030 sonar~\cite{Hansen,Callow,et.al}, and on simulations of the same sonar with 100 realisations of speckle. The simulations contained a few highlighted cylindric regions with 2\,m diameter and a constant intensity 15.4\,dB over the average speckle level. We focus on an object centered at 41\,m range and -3 degrees azimuth angle. The images produced for a single speckle realisation, and the mean image of all 100 realisations in \Fig{speckle}. Each image was normalised by their respective speckle level, and the dynamic range was clamped at \{-30, 15\}\,dB. The physical extent of the of the simulated object is superimposed for reference. %\ref{saf09}

We see that both the adaptive beamformers have far better angular sidelobe suppression than DAS, and produce a more clearly defined shadow. This is also observed in \fig{cutmean}, which displays a cut through the shadow and highlight region of the image. The transition region between highlight and shadow is shortest for the LCA, implying that its angular resolution is better than that of the MV beamformer. This is advantageous since it causes less leakage of highlight into shadow areas, thereby resulting in a more accurate representation of the object size. The slightly inferior MV resolution is due to the subarray averaging, which makes the effective array smaller.

We have found that the tendency of the LCA to perform similarly to the MV beamformer is consistent in most scenarios, except when it is of essence that strong interferers close to object being imaged are present. In such cases the MV beamformer is expected to take advantage of its full solution space to design a spatial response with zero sensitivity at the interferers angles, while LCA will be limted to the spatial responses that we selected. In short, whenever a scenario comes along that favors a spatial response we did not think of, the MV will triumph. On the other hand, increasing the LCA solution space will at some point leave it with nearly the same freedom as the MV, without suffering from statistical instability.

By studying the LCA performance on the 100 speckle realisations, we found the choice of window to be highly dependent on the phenomena being imaged. In speckle regions LCA favored narrow spatial responses with minor steering, while in shadow regions a wide and fully steered responses were often selected. There was also a clear distinction between the responses selected in sidelobe regions and in speckle regions.

@@

In \Fig{holmengraa} the wreck Holmengraa is imaged using the LCA and DAS beamformer. It is 26\,m long and 6\,m wide, and lies on a slanted seabed at 77\, depth. Again, the LCA produces a cleaner shadow and better edge definition. The MV method performed almost identically to LCA in this case, and was omitted.

@@

The LCA method delivers performance similar to the MV method, but at a fraction of the computational cost. This is verified in our simulations of speckle, and on experimental data.If the windows are With a careful window design, If the window database is designed wisely,  a window database to suit

% \printreferences
% If bibtex:
\bibliography{references2}
